Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-29 TACAS2018/paper.tex
nudge

Let $\mathbb{R}_{\langle M,F \rangle}$ be a fixed point domain with $M$ bits
representing the binary word length of the representation and $F$ bits
representing the pre-scaling $2^{-F}$ (\emph{i.e} the number of bits after
the decimal place).  We note that a negative $F$ denotes an integer
multiplication that adds $-F$ trailing zeros in the number represented by
$M$. 
We will use the domain $\mathbb{R}_{\langle C \rangle} \text{ where } \langle C \rangle = \langle M_c,F_c \rangle$
to represent the fixed point precision of the controller, and 
$\mathbb{R}_{\langle P \rangle} \supseteq \mathbb{R}_{\langle C \rangle}$
to represent the domain of a digital model of the dynamics of a real plant when
evaluated by a verification tool. \note{refer back to notation in Figure 1.}
Thus, any mathematical operation within the digital controller is executed in the
domain $\mathbb{R}_{\langle C \rangle}$, and all other calculations
in our model will be carried out in the domain $\mathbb{R}_{\langle P \rangle}$.
Given ${\langle M,F \rangle}$, 
we denote as $\mathcal{F}_{\langle M,F \rangle} : \mathbb{R} \rightarrow \mathbb{R}_{\langle M,F \rangle}$ the corresponding truncation function.

@@

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{order} & \multicolumn{1}{|c|}{Number of} & \multicolumn{3}{|c|}{CEGIS} & \multicolumn{3}{|c|}{CEGIS-OR} \\
   &                                 & & \multicolumn{1}{|c|}{benchmarks}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{Max} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{Max}\\\hline
1  & DC Motor          & 2 & 100 & 1.3\,s & 26.9\,s& \xmark (22)& 1.1\,s &   3.8\,s &   \xmark (2)\\
2  & Helicopter        & 3  & 100 & 57.2\,s & \xmark (54) & \xmark (54) & 7.1\,s  &   18.0\,s & \xmark (9)\\
3  & Inverted Pendulum & 4 & 100 &   9.7\,s & \xmark &\xmark () & 12.0\,s  &  18.1\,s &  \xmark ()\\
4  & Magnetic Pointer  & 2  & 100 & 5.7\,s & 229.9\,s & \xmark (38) & 3.9\,s  &  11.4\,s &  \xmark (5) \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 14.0\,s & 2.0\,s & \xmark ()&0.7\,s  &   6.3\,s & \xmark  (3)\\
6  & Suspension        & 4 & 100 & 73.7\,s& \xmark & \xmark ()&24.0\,s  &   46.1\,s & \xmark ()\\
7  & Tape Driver       & 3 & 100 & 10.1\,s& 140.3\,s & \xmark ()&2.6\,s  &   10.5\,s & \xmark ()\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results. Minimum, Maximum and Median times for standard CEGIS vs CEGIS-OR on multiple benchmarks. \xmark~indicates timeout, with the number of benchmarks that timed out in parenthesis.}
\end{table}

Using the described technique we managed to synthesize controllers for X benchmarks, with a median synthesis time of $18.8$ seconds.
%and the maximum observed synthesis time was under 10 minutes for a $4^{th}$ order system ($8^{th}$ order considering the observer).
We consider these times short enough to be of practical use to control engineers.
There are cases for which the system fails to find a controller.  We use a
relatively short timeout, so it is possible longer timeouts would result in a
valid solution. However, as stated before, the solution is sound but incomplete,
so it is impossible to tell if a solution exists in these cases.

In order to show the improvements due to our optimization-enhanced CEGIS, we
compared our approach against a standard CEGIS procedure.  For the majority of
our benchmarks, we observe that our approach is significantly faster than
standard CEGIS. We further note the cases where standard CEGIS times out, which
represent an improvement of several orders of magnitude for our optimization
approach.  The magnitude of the improvement is largely dependent on the shape
of the problem.

The smaller benchmarks (order=2) tend to show little or no improvement when we
apply CEGIS-OR.  The reason is that for these benchmarks the synthesis time using
Solution Generalization is already small, and the overhead of the abstraction is
often equal or larger than the gain obtained.

>>>>

Let $\mathbb{R}_{\langle M,F \rangle}$ be a fixed point domain with $M$ bits
representing the binary word length of the representation and $F$ bits
representing the pre-scaling $2^{-F}$ (\emph{i.e} the number of bits after
the decimal place).  We note that a negative $F$ denotes an integer
multiplication that adds $-F$ trailing zeros in the number represented by
$M$. 
We will use the domain $\mathbb{R}_{\langle C \rangle} \text{ where } \langle C \rangle = \langle M_c,F_c \rangle$
to represent the fixed point precision of the controller, and 
$\mathbb{R}_{\langle P \rangle} \supseteq \mathbb{R}_{\langle C \rangle}$
to represent the domain of a digital model of the dynamics of a real plant when
evaluated by a verification tool.  Elements in this domain are marked with an
underscore~${\langle C \rangle}$ (see Figure \ref{fig:observersystem2}).
Thus, any mathematical operation within the digital controller is executed in the
domain $\mathbb{R}_{\langle C \rangle}$, and all other calculations
in our model will be carried out in the domain $\mathbb{R}_{\langle P \rangle}$.
Given ${\langle M,F \rangle}$, 
we denote as $\mathcal{F}_{\langle M,F \rangle} : \mathbb{R} \rightarrow \mathbb{R}_{\langle M,F \rangle}$ the corresponding truncation function.

@@

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{order} & \multicolumn{1}{|c|}{Number of} & \multicolumn{3}{|c|}{CEGIS} & \multicolumn{3}{|c|}{CEGIS-OR} \\
   &                                 & & \multicolumn{1}{|c|}{benchmarks}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{Max} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{Max}\\\hline
1  & DC Motor          & 2 & 100 & 1.3\,s & 26.9\,s& \xmark (22)& 1.1\,s &   3.8\,s &   \xmark (2)\\
2  & Helicopter        & 3  & 100 & 57.2\,s & \xmark (54) & \xmark (54) & 7.1\,s  &   18.0\,s & \xmark (9)\\
3  & Inverted Pendulum & 4 & 100 &   9.7\,s & \xmark &\xmark () & 12.0\,s  &  18.1\,s &  \xmark ()\\
4  & Magnetic Pointer  & 2  & 100 & 5.7\,s & 229.9\,s & \xmark (38) & 3.9\,s  &  11.4\,s &  \xmark (5) \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 14.0\,s & 2.0\,s & \xmark ()&0.4\,s  &   4.9\,s & \xmark  (6)\\
6  & Suspension        & 4 & 100 & 73.7\,s& \xmark & \xmark ()&24.0\,s  &   46.1\,s & \xmark ()\\
7  & Tape Driver       & 3 & 100 & 10.1\,s& 140.3\,s & \xmark ()&2.6\,s  &   10.5\,s & \xmark ()\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results. Minimum, Maximum and Median times for standard CEGIS vs CEGIS-OR on multiple benchmarks. \xmark~indicates timeout, with the number of benchmarks that timed out in parenthesis.}
\end{table}

Using the described technique we managed to synthesize controllers for X benchmarks, with a median synthesis time of $18.8$ seconds.
%and the maximum observed synthesis time was under 10 minutes for a $4^{th}$ order system ($8^{th}$ order considering the observer).
We consider these times short enough to be of practical use to control engineers.
There are cases for which the system fails to find a controller.  We use a
relatively short timeout, so it is possible that longer timeouts would result in a
valid solution. However, as stated before, the solution is sound but incomplete,
so it is impossible to tell if a solution exists in these cases.

In order to show the improvements due to our optimization-enhanced CEGIS, we
compared our approach against a standard CEGIS procedure.  For the majority of
our benchmarks, we observe that our approach is significantly faster than
standard CEGIS. We further note that the number of times standard CEGIS times
out is much larger than CEGIS-OR which results in more found solutions.
The magnitude of the improvement is however, largely dependent on the shape of the problem.
